---
title: "In-Class Exercise 7: Global Measures of Spatial Autocorrelation"
date: "20 February 2023"
date-modified: "`r Sys.Date()`"
format: html
execute:
  eval: true
  echo: true
  message: false
  warning: false
editor: visual
---

# 1 Getting Started

## 1.1 Importing Modules

```{r}
pacman::p_load(sf, sfdep, tmap, tidyverse, plotly)
```

## 1.2 Importing Data

### 1.2.1 Shapefile

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### 1.2.2 CSV

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

### 1.2.3 Regional Join

```{r}
hunan_GDPPC <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)
```

### 1.2.4 Plotting a Choropleth Map

```{r}
tmap_mode("plot")
tm_shape(hunan_GDPPC) +
  tm_fill("GDPPC", 
          style = "quantile", 
          palette = "Blues",
          title = "GDPPC") +
  tm_layout(main.title = "Distribution of GDP per capita by district, Hunan Province",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2)
```

# 2 Global Measures of Spatial Association

## 2.1 Deriving Contiguity Weights: Queen's Method

```{r}
wm_q <- hunan_GDPPC %>%   
  mutate(nb = st_contiguity(geometry),          wt = st_weights(nb,                          style = "W"),          .before = 1)
```

## 2.2 Computing Global Moran's I

```{r}
moranI <- global_moran(wm_q$GDPPC,
                       wm_q$nb,
                       wm_q$wt)
```

## 2.3 Performing Global Moran's I Test

```{r}
global_moran_test(wm_q$GDPPC,
                  wm_q$nb,
                  wm_q$wt)
```

## 2.4 Performing Global Moran's I Permutation Test

```{r}
set.seed(1234)
global_moran_perm(wm_q$GDPPC,
                 wm_q$nb,
                 wm_q$wt,
                 nsim = 99)
```

# 3 Computing Local Moran's I

## 3.1 Computing Local Moran's I

```{r}
lisa <- wm_q %>%
  mutate(local_moran = local_moran(
    GDPPC, nb, wt, nsim = 99),
      .before = 1) %>%
  unnest(local_moran)
  #important to unnest
lisa
```

## 3.2 Visualising Local Moran's I

```{r}
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("ii") +
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6, 8))
```

## 3.3 Visualising p-value of Moran's I

```{r}
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("p_ii") +
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6, 8))
```

## 3.4 Visualising Local Moran's I

```{r}
lisa_sig <- lisa %>%
  filter(p_ii < 0.05)
tmap_mode("plot")
tm_shape(lisa) +
  tm_polygons() +
  tm_borders(alpha = 0.5) +
tm_shape(lisa_sig) +
  tm_fill("mean") +
  tm_borders(alpha = 0.4)
```

# 4 Hot Spot and Cold Spot Area Analysis

## 4.1 Computing Local Moran's I

```{r}
HCSA <- wm_q %>%
  mutate(local_Gi = local_gstar_perm(
    GDPPC, nb, wt, nsim = 99),
    .before = 1) %>%
  unnest(local_Gi)
HCSA
```

## 4.2 Visualising Gi\*

```{r}
tmap_mode("view")
tm_shape(HCSA) +
  tm_fill("gi_star") +
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6, 8))
```

## 4.3 Visualising p-value of HCSA

```{r}
tmap_mode("plot")
tm_shape(HCSA) +
  tm_fill("p_sim") +
  tm_borders(alpha = 0.5)
```

# EHSA

# 1 Getting Started

## 1.1 Importing Module

```{r}
pacman::p_load(sf, sfdep, tmap, tidyverse, plotly)
```

## 1.2 Importing Data

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

```{r}
GDPPC <- read_csv("data/aspatial/Hunan_GDPPC.csv")
```

```{r}
GDPPC_st <- spacetime(GDPPC, hunan,
                      .loc_col = "Country",
                      .time_col = "Year")
```

```{r}
GDPPC_nb <- GDPPC_st %>%
  activate("geometry") %>%
  mutate(
    nb = include_self(st_contiguity(geometry)),
    wt = st_weights(nb)
  ) %>%
  set_nbs("nb") %>%
  set_wts("wt")
```

```{r}
gi_stars <- GDPPC_nb %>%
  group_by(Year) %>%
  mutate(gi_star = local_gstar_perm(
    GDPPC, nb, wt, nsim = 99)) %>%
  tidyr::unnest(gi_star)
```

```{r}

```
