---
title: "Hands-On Exercise 2: Geospatial Data Wrangling"
date: "30 January 2023"
date-modified: "`r Sys.Date()`"
format: html
execute:
  message: false
  warning: false
editor: visual
---

# Getting Started

## Importing Modules

```{r}
pacman::p_load(sf, tidyverse)
```

## Importing Data

### Geospatial Data

#### Importing Polygon Feature Data in shapefile Format

```{r}
mpsz = st_read(dsn = "data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

#### Importing Polyline Feature Data in shapefile Form

```{r}
cyclingpath = st_read(dsn = "data/geospatial", 
                         layer = "CyclingPathGazette")
```

#### Importing GIS Data in kml Format

```{r}
preschool = st_read("data/geospatial/preschools-location.kml")
```

# Checking the Content of a Simple Feature Data Frame

## Working with st_geometry()

```{r}
st_geometry(mpsz)
```

## Working with glimpse()

```{r}
glimpse(mpsz)
```

## Working with head()

```{r}
head(mpsz, n=5)  
```

# Plotting the Geospatial Data

```{r}
plot(mpsz)
```

```{r}
plot(st_geometry(mpsz))
```

```{r}
plot(mpsz["PLN_AREA_N"])
```

# Working with Projection

## Assigning EPSG Code to a Simple Feature Data Frame

```{r}
st_crs(mpsz)
```

```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

```{r}
st_crs(mpsz3414)
```

## Transforming the Projection of Preschool from WGS84 to SVY21

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

# Importing and Converting an Aspatial Data

## Importing the Aspatial Data

```{r}
listings <- read_csv("data/aspatial/listings.csv")
```

## Creating a Simple Feature Data Frame From an Aspatial Data Frame

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

```{r}
glimpse(listings_sf)
```

# Geoprocessing with sf Package

## Buffering

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
```

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
```

```{r}
sum(buffer_cycling$AREA)
```

## Point-in-Polygon Count

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

```{r}
summary(mpsz3414$`PreSch Count`)
```

```{r}
top_n(mpsz3414, 1, `PreSch Count`)
```

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
```

```{r}
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

# Explorotary Data Analysis (EDA)

```{r}
hist(mpsz3414$`PreSch Density`)
```

```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

```{r}
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
